Embedding Intuitionistic into Classical Logic

### 2023-03-09 LPAR/LPAR-EIICL.tex
formatting

For predicate logic we assume an ambient signature $\Sigma$, i.e., a set of function and predicate symbols with their respective arities, and a set of variables. As usual we can inductively define terms and formuas. A term is either a variable or, given an $n$-ary function symbol $f$ and terms $t_1, \ldots, t_n$, $f(t_1, \ldots, t_n)$. A formula is either $\bot$, or, given a $n$-ary predicate symbol $P$ and terms $t_1, \ldots, t_n$, $P(t_1, \ldots, t_n)$, or, given formulas $\varphi$ and $\psi$ and $\circ\in\{\wedge, \vee,\to\}$, $\varphi\circ\psi$, or, given a formula $\varphi$ and a variable $x$, $\exists x\varphi$ or $\forall x\varphi$. Free and bound variables are defined as usual and we assume that they comprise distinct sets, such that bound variables are not affected by substitutions of free variables. Generally we will denote free variables with letters at the beginning of the alphabet and bound variables with letters at the end of the alphabet.

@@

An important step in the embedding will be the elimination of quantifiers via Herbrandization.
In this process we introduce fresh variables and add additional function symbols to the signature.
A fresh variable is any variable that does not occur in any of the considered formulas.
Whenever we add a function symbol we implicitly extend the signature by some not previously contained symbol.

\begin{definition}\label{definition:skolemization}
	For formulas $\varphi$ we define the Skolemization $\varphi^S_Z$ and Herbrandization $\varphi^H_Z$ with respect to a set of free variables $Z$ by simultaneous induction:
	\begin{itemize}
		\item $A^S_Z = A^H_Z = A$ for each atomic $A$.
		\item $(\varphi\circ\psi)^X_Z = \varphi^X_Z\circ\psi^X_Z$ for $\circ\in\{\wedge, \vee\}$, $X\in\{S, H\}$.
		\item $(\varphi\to\psi)^S_Z = \varphi^H_Z\to \psi^S_Z$ and $(\varphi\to\psi)^H_Z = \varphi^S_Z\to\psi^H_Z$.
		\item $(\forall x\varphi)^S_Z = \forall x(\varphi[a/x]^S_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.\\$(\forall x\varphi)^H_Z = \varphi[s(z_1,\dots,z_n)/x]^H_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.
		\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.\\$(\exists x\varphi)^H_Z = \exists x(\varphi[a/x]^H_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.
	\end{itemize}
	Let $$\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$$ and $$\varphi^H = (\forall x_1\dots\forall x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
\end{definition}

@@

A similar result holds in the first order case.

@@

Since we could not find a proof of this, one is given in the appendix.

@@

The general idea is to use the Kripke semantics for intuitionistic logic, which are expressed in classical logic, to perform an embedding of formulas. That is, for each formula $\varphi$ we give a formula $\varphi^\#$ that is classically valid if and only if $\varphi$ is intuitionistically valid, establishing a dual to the double negation translation.

@@

This is somewhat unsatisfying as we end up in a much more complex logic. However using the normal form from Lemma~\ref{lemma:propositional-normal-form}, i.e. assuming that $\varphi$ is of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$ for some flat clauses $\mathcal R$ and implicational clauses $\mathcal X$, further simplification is possible:
\begin{itemize}
	\item For flat clauses $(\bigwedge A_i\to\bigvee B_j)^\nu$ is
	$$\forall w\left(\nu\preceq w\to\bigwedge_i A_i(w)\to\bigvee_jB_j(w)\right).$$
	\item For implication clauses $((A\to B)\to C)^\nu$ is
	$$\forall w(\nu\preceq w\to(\forall k(w\preceq k\to A(k)\to B(k)))\to C(w))$$
\end{itemize}

and after applying Herbrandization we end up with a formula $$\varphi^{CH} = K(\varphi)\to\left(\bigwedge\mathcal R^{b S}\wedge \bigwedge\mathcal X^{b S}\right)\to P(s)$$ where $s$ is a new constant symbol, each formula in $\mathcal R^{b S}$ is of the form
$$\forall w\left(s\preceq w\to\bigwedge_iA_i(w)\to\bigvee_jB_j(w)\right)$$
and each formula $\psi\in\mathcal X^{b S}$ is of the form
$$\forall w(s\preceq w\to(w\preceq f_\psi(w)\to A(f_\psi(w))\to B(f_\psi(w)))\to C(w))$$
with $f_\psi$ being a new function symbol.

We now argue that if a counter-model for $\varphi^{CH}$ exists then a small counter-model exists. This is ultimately what enables our translations.

@@

\begin{corollary}
	If a formula in normal form as in Lemma~\ref{lemma:propositional-normal-form} is intuitionistically invalid then there exists a Kripke counter-model which is a rooted tree with height and degree bounded by $|\mathcal X|$.
\end{corollary}

@@

We now give an analogous transformation for first order logic. Our trick of lifting the sentence to first-order logic to encode the Kripke semantics no longer works in such a straightforward manner, since we already are in first-order logic. However we do it nonetheless! The domain of our classical model will feature on the one hand elements representing worlds in a Kripke frame and on the other hand the domain-elements from the Kripke model. We reconcile these notions by introducing a special binary predicate $E$, first considered in~\cite{baaz2006skolemization} and expanded onto in~\cite{iemhoff2010eskolemization}, where $E(x, u)$ encodes that $x$ is an element of the domain of the world $u$. To encode that some $n$-ary predicate $A$ holds at a world $u$ we extend each $n$-ary predicate symbol to a $n+1$-ary predicate symbol of the same name, interpreting $A(\vec x, u)$ as "$A(\vec x)$ holds at u". Furthermore we write $\vec E(\vec x, u)$ for $\bigwedge\{E(x_i, u)\:|\:x_i\in \vec x\}$. With this we obtain the following encoding of Kripke Semantics:

\begin{definition}\label{def:fo-translation}
	Let $\varphi$ be a predicate formula. Define $\varphi^{\nu}$ inductively:
	\begin{itemize}
		\item $\bot^\nu = \bot$.
		\item $A(\vec t)^{\nu} = A(\vec t, \nu)$ for every predicate symbol $A$.
		\item $(\varphi\circ\psi)^\nu = \varphi^\nu\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
		\item $(\varphi\to \psi)^\nu = \forall w(\nu\preceq w\to\varphi^{w}\to\psi^{w})$.
		\item $(\exists x\varphi)^\nu = \exists x(E(x,\nu)\wedge\varphi[a/x]^\nu[x/a])$ where $a$ is a new free variable.
		\item $(\forall x\varphi)^\nu = \forall w(i\preceq w\to \forall x(E(x, w)\to \varphi[a/x]^w[x/a]))$ where $a$ is a new free variable and $w$ is a new bound variable.
	\end{itemize}
	where $w$ is some new bound variable. Let $K(\varphi)$ encode the theory of Kripke structures as well as the predicate $E$, i.e.
	\begin{align*}
		K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
		& \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge \forall u\forall w (u\preceq w\to \text{Persistent}(u, w)).
	\end{align*}
	with e.g.:
	\begin{align*}
		\text{PartialOrder}(\preceq) = &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
		\text{DomainSubset}(u, w) = &\:\forall x(E(x, u)\to E(x, w))\\
		\text{World}(u) = &\:\exists xE(x, u)\\
		\text{DomainClosed}(u) = &\:\bigwedge\left\{\forall\vec x(\vec E(\vec x, u)\to E(f(\vec x), u))\:|\:\text{$f$ a function in $\varphi$}\right\}\\
		\text{Persistent}(u, w) = &\:\bigwedge\left\{\forall\vec x(\vec A(\vec x, u)\to A(\vec x, w))\:|\:\text{$A$ a predicate in $\varphi$}\right\}
	\end{align*}
	Then define
	$$\varphi^C = K(\varphi)\to\text{World}(b)\to \vec E(\vec a, b)\to \varphi^b$$where $\vec a$ contains all free variables occurring in $\varphi$ and $b$ is a new free variable.
\end{definition}

@@

As in the previous section we can define a tree counter-model for this.
\begin{definition}
	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ are the sequences on $\{(\vec m, \psi)\:|\:\psi\in \mathcal F, \vec m\in M^n\}$.
		\item $r^{I_T} = s^{I_T} = \epsilon$.
		\item We interpret $\preceq$ as the prefix-order.
		\item We set $$f_\psi^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)(\vec m, \psi).$$
		\item For predicate symbols $R$, we set $${R}^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = {R}^I(\vec m, f_{\psi_n}(\vec m_n, \dots(f_{\psi_1}(\vec m_1, s^I))\dots)).$$
	\end{itemize}
\end{definition}

@@

Even if we don't have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$f_{\psi_1}(\vec m_1, f_{\psi_2}(\vec m_2, \dots f_{\psi_n}(\vec m_n, s)\dots))$$ is sufficient.
In particular, this allows us to completely eliminate $\preceq$, that is we remove the PartialOrder$(\preceq)$
from $K(\varphi)$ and adjust DomainSubset$(u, w)$ and Persistent$(u, w)$, e.g. remove quantification over $w$ and the precondition $u\preceq w$ from $K(\varphi)$ and replace DomainSubset$(u, w)$ with
$$
	\bigwedge\{\forall u\forall x\forall\vec y(E(x, u)\to E(x, f_\psi(\vec y, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q\}
$$
and Persistent$(u, w)$ with
$$
	\bigwedge\{\forall\vec x\forall\vec y\forall u(A(\vec x, u)\to A(\vec x, f_\psi(\vec y, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q, \text{$A$ a predicate in $\varphi$}\}.
$$Putting all of this together yields the following translation:
\begin{theorem}\label{thm:fo-translation}
	Let $\mathcal R,\mathcal X,\mathcal Q$ be as above. For every $n$-ary predicate $A$ define a new $n+1$-ary predicate of the same name. For every $\psi\in\mathcal \mathcal X\cup\mathcal Q$ define a new function symbol $f_\psi$. Consider a new binary predicate symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$. Obtain $\mathcal R^\#$ by including
	$$
		\forall u\forall \vec x\left(\vec E(\vec x, u)\to\bigwedge_i A_i(\vec a_i, u)\to\bigvee_j B_j(\vec b_j, u)\right)
	$$
 	for each flat clause
 	$$
 		\forall \vec x\left(\bigwedge_i A_i(\vec a_i)\to\bigvee_j B_j(\vec b_j)\right)\in\mathcal R,
 	$$
 	$\mathcal X^\#$ by including
 	$$
 		\forall u\forall \vec x\left(A(\vec a, f_\psi(\vec x, u))\to B(\vec b, f_\psi(\vec x, u))\to C(\vec c, u)\right)
 	$$
 	for each implication clause
	$$
		\psi = \forall \vec x\left(\left(A(\vec a)\to B(\vec b)\right)\to C(\vec c)\right)\in\mathcal X,
	$$
	and $\mathcal Q^\#$ by including
	$$
		\forall u\forall \vec x\left(\forall y\left(A(\vec a, f_\psi(\vec x, u))\right)\to B(\vec b, u)\right)
	$$
	for each implication clause
	$$
		\psi = \forall \vec x\left(\forall y\left(A(\vec a)\right)\to B(\vec b)\right)\in\mathcal X.
	$$
	Then $\varphi$ is intuitionistically valid if and only if
	$$
		\varphi^\# := \exists xE(x, s)\to K(\varphi)\to \left(\bigwedge\mathcal R^\#\wedge\bigwedge\mathcal X^\#\wedge\bigwedge\mathcal Q^\#\right)\to P^\#(s)
	$$
	is classically valid, where $s$ is a new constant symbol and
	\begin{align*}
		K(\varphi) := &\:\bigwedge\left\{\forall\vec z\forall x\forall u((E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q\right\}\wedge\\
		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ a function in $\varphi$}\}\right)\wedge\\
		&\resizebox{.9\hsize}{!}{$\:\bigwedge\left\{\forall\vec z_1\forall \vec z_2\forall u(A(\vec z_1, u)\to A(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal X\cup\mathcal Q$, $A$ a predicate in $\varphi$}\right\}$.}
	\end{align*}
	The size of the obtained formula is linear in the input. However $|\mathcal X| + |\mathcal Q| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary predicate has been extended to a $n+1$-ary predicate.
\end{theorem}

It is not possible to completely eliminate worlds from the domain by enumerating them since there are formulas with no finite counter-examples. An analogous procedure to the propositional case fails, however we can still retrieve additional axioms that restrict the interpretation of functions. That is, we may assume that $f_\psi(\vec t, w) = w$, if $\exists u(w = f_\psi(\vec t, u))$ for all $\psi\in\mathcal X\cup\mathcal Q$ (i.e. that $f_\psi$ parametrized by $\vec t$ is idempotent), and $f_\psi(\vec t, w) = w$, if $\exists u(w \geq f_\psi(\vec t, u))$ for all $\psi\in\mathcal X$ (which is even stronger than idempotence).

@@

\begin{wrapfigure}[4]{r}{0.4\textwidth}
	\vspace*{-2cm}
	\begin{tabular}{l|c|c}
		Embedding&571&21.4\%\\
		Embedding (idem.)&557&20.9\%\\
		ileanCoP 1.2&875&32.8\%\\
		nanoCoP-i 2.0&858&32.1\%\\\hline
		Total&2669&100\%
	\end{tabular}
	\caption{Solved problems.}
	\label{fig:a}
\end{wrapfigure}

@@

\begin{wrapfigure}[16]{r}{0.4\textwidth}
	\vspace{6cm}
	\hspace*{1.5cm}
	\begin{tikzpicture}
		\begin{scope}[opacity=0.1, blend mode=multiply]
			\def\firstcircle{(90:.75) circle (1.5)}
			\def\secondcircle{(210:.75) circle (1.5)}
			\def\thirdcircle{(330:.75) circle (1.5)}
			\begin{tikzpicture}
				\begin{scope}[shift={(3cm,-5cm)}, fill opacity=0.5]
					\fill[red] \firstcircle;
					\fill[green] \secondcircle;
					\fill[blue] \thirdcircle;
					\draw node {$491$};
					\draw node at (30:1.2) {$30$};
					\draw node at (150:1.2) {$42$};
					\draw node at (270:1.2) {$22$};
					\draw node at (90:1.8) {$8$};
					\draw node at (210:1.8) {$2$};
					\draw node at (330:1.8) {$417$};
				\end{scope}
			\end{tikzpicture}
		\end{scope}
	\end{tikzpicture}
	\vspace*{-7.5cm}
	\caption{Instances proven by combinations of Embedding (red), Embedding (idem.) (green), ileanCoP or nanoCoP (blue)}
	\label{fig:c}
\end{wrapfigure}

>>>>

For predicate logic we assume an ambient signature $\Sigma$, i.e., a set of function and predicate symbols with their respective arities, and a set of variables. A term is either a variable or, given an $n$-ary function symbol $f$ and terms $t_1, \ldots, t_n$, $f(t_1, \ldots, t_n)$. A formula is either $\bot$, or, given a $n$-ary predicate symbol $P$ and terms $t_1, \ldots, t_n$, $P(t_1, \ldots, t_n)$, or, given formulas $\varphi$ and $\psi$ and $\circ\in\{\wedge, \vee,\to\}$, $\varphi\circ\psi$, or, given a formula $\varphi$ and a variable $x$, $\exists x\varphi$ or $\forall x\varphi$. Free and bound variables are defined as usual and we assume that they comprise distinct sets, such that bound variables are not affected by substitutions of free variables. Generally letters at the beginning of the alphabet will denote free variables and letters at the end of the alphabet bound variables.

@@

An important step in the embedding will be the elimination of quantifiers via Herbrandization.
In this process we introduce fresh variables and add additional function symbols to the signature.
A fresh variable is any variable that does not occur in any considered formula.
Whenever we add a function symbol we implicitly extend the signature by a not previously contained symbol.

\begin{definition}\label{definition:skolemization}
	For formulas $\varphi$ we define the Skolemization $\varphi^S_Z$ and Herbrandization $\varphi^H_Z$ with respect to a set of free variables $Z$ by simultaneous induction:
	\begin{itemize}
		\item $A^S_Z = A^H_Z = A$ for each atomic $A$.
		\item $(\varphi\circ\psi)^X_Z = \varphi^X_Z\circ\psi^X_Z$ for $\circ\in\{\wedge, \vee\}$, $X\in\{S, H\}$.
		\item $(\varphi\to\psi)^S_Z = \varphi^H_Z\to \psi^S_Z$ and $(\varphi\to\psi)^H_Z = \varphi^S_Z\to\psi^H_Z$.
		\item $(\forall x\varphi)^S_Z = \forall x(\varphi[a/x]^S_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.\\$(\forall x\varphi)^H_Z = \varphi[s(z_1,\dots,z_n)/x]^H_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.
		\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.\\$(\exists x\varphi)^H_Z = \exists x(\varphi[a/x]^H_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.
	\end{itemize}
	Let $\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$ and $\varphi^H = (\forall x_1\dots\forall x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
\end{definition}

@@

@@

Since we could not find a proof of this, one is given in the appendix~\ref{proof:first-order-normal-form}.

@@

The idea is to use the Kripke semantics for intuitionistic logic, expressed in classical logic, to perform an embedding of formulas. That is, for each $\varphi$ we give a $\varphi^\#$ that is classically valid iff $\varphi$ is intuitionistically valid, establishing a dual to the double negation translation.

@@

This is somewhat unsatisfying as we end up in a much more complex logic. However using the normal form from Lemma~\ref{lemma:propositional-normal-form}, i.e. assuming that $\varphi$ is of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$ for some flat clauses $\mathcal R$ and implicational clauses $\mathcal X$, further simplification is possible:
\begin{itemize}
	\item For flat clauses $(\bigwedge A_i\to\bigvee B_j)^\nu$ is
	$\forall w\left(\nu\preceq w\to\bigwedge_i A_i(w)\to\bigvee_jB_j(w)\right).$
	\item For implication clauses $((A\to B)\to C)^\nu$ is
	$$\forall w(\nu\preceq w\to(\forall k(w\preceq k\to A(k)\to B(k)))\to C(w))$$
\end{itemize}

and after applying Herbrandization we end up with a formula $$\varphi^{CH} = K(\varphi)\to\left(\bigwedge\mathcal R^{b S}\wedge \bigwedge\mathcal X^{b S}\right)\to P(s)$$ where $s$ is a new constant symbol, each formula in $\mathcal R^{b S}$ is of the form
$$\forall w\left(s\preceq w\to\bigwedge_iA_i(w)\to\bigvee_jB_j(w)\right)$$
and each formula $\psi\in\mathcal X^{b S}$ is of the form
$$\forall w(s\preceq w\to(w\preceq f_\psi(w)\to A(f_\psi(w))\to B(f_\psi(w)))\to C(w))$$
with $f_\psi$ being a new function symbol.
We now argue that if a counter-model for $\varphi^{CH}$ exists then a small counter-model exists. This is ultimately what enables our translations.

@@

\begin{corollary}
	If a formula in normal form as in Lemma~\ref{lemma:propositional-normal-form} is intuitionistically invalid then there exists a Kripke counter-model that is a rooted tree of height and degree bounded by $|\mathcal X|$.
\end{corollary}

@@

We now give an analogous transformation for first order logic. Our trick of lifting the sentence to first-order logic to encode the Kripke semantics no longer works in such a straightforward manner, since we already are in first-order logic. However we do it nonetheless! The domain of our classical model will feature on the one hand elements representing worlds in a Kripke frame and on the other hand the domain-elements from the Kripke model. We reconcile these notions by introducing a special binary predicate $E$, first considered in~\cite{baaz2006skolemization} and expanded onto in~\cite{iemhoff2010eskolemization}, where $E(x, u)$ encodes that $x$ is an element of the domain of the world $u$. To encode that some $n$-ary predicate $A$ holds at a world $u$ we extend each $n$-ary predicate symbol to a $n+1$-ary predicate symbol of the same name, interpreting $A(\vec x, u)$ as "$A(\vec x)$ holds at u". Write $\vec E(\vec x, u)$ for $\bigwedge\{E(x_i, u)\:|\:x_i\in \vec x\}$. With this we obtain the following encoding of Kripke Semantics:

\begin{definition}\label{def:fo-translation}
	Let $\varphi$ be a predicate formula. Define $\varphi^{\nu}$ inductively:
	\begin{itemize}
		\item $\bot^\nu = \bot$.
		\item $A(\vec t)^{\nu} = A(\vec t, \nu)$ for every predicate symbol $A$.
		\item $(\varphi\circ\psi)^\nu = \varphi^\nu\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
		\item $(\varphi\to \psi)^\nu = \forall w(\nu\preceq w\to\varphi^{w}\to\psi^{w})$.
		\item $(\exists x\varphi)^\nu = \exists x(E(x,\nu)\wedge\varphi[a/x]^\nu[x/a])$ where $a$ is a new free variable.
		\item $(\forall x\varphi)^\nu = \forall w(i\preceq w\to \forall x(E(x, w)\to \varphi[a/x]^w[x/a]))$ where $a$ is a new free variable and $w$ is a new bound variable.
	\end{itemize}
	where $w$ is some new bound variable. Let $K(\varphi)$ encode the theory of Kripke structures as well as the predicate $E$, i.e.
	\begin{align*}
		K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
		& \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge \forall u\forall w (u\preceq w\to \text{Persistent}(u, w)).
	\end{align*}
	with e.g.:\vspace*{-.76cm}
	\begin{align*}
		\text{PartialOrder}(\preceq) = &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
		\text{DomainSubset}(u, w) = &\:\forall x(E(x, u)\to E(x, w))\\
		\text{World}(u) = &\:\exists xE(x, u)\\
		\text{DomainClosed}(u) = &\:\bigwedge\left\{\forall\vec x(\vec E(\vec x, u)\to E(f(\vec x), u))\:|\:\text{$f$ a function in $\varphi$}\right\}\\
		\text{Persistent}(u, w) = &\:\bigwedge\left\{\forall\vec x(\vec A(\vec x, u)\to A(\vec x, w))\:|\:\text{$A$ a predicate in $\varphi$}\right\}
	\end{align*}
	Then define
	$\varphi^C = K(\varphi)\to\text{World}(b)\to \vec E(\vec a, b)\to \varphi^b$ where $\vec a$ contains all free variables occurring in $\varphi$ and $b$ is a new free variable.
\end{definition}

@@

As in the previous section we can define a tree counter-model for this.
\begin{definition}
	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ are the sequences on $\{(\vec m, \psi)\:|\:\psi\in \mathcal F, \vec m\in M^n\}$.
		\item $r^{I_T} = s^{I_T} = \epsilon$.
		\item We interpret $\preceq$ as the prefix-order.
		\item We set $f_\psi^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)(\vec m, \psi).$
		\item For predicate symbols $R$, we set $${R}^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = {R}^I(\vec m, f_{\psi_n}(\vec m_n, \dots(f_{\psi_1}(\vec m_1, s^I))\dots)).$$
	\end{itemize}
\end{definition}

@@

Even if we don't have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$f_{\psi_1}(\vec m_1, f_{\psi_2}(\vec m_2, \dots f_{\psi_n}(\vec m_n, s)\dots))$$ is sufficient.
In particular, this allows us to completely eliminate $\preceq$, that is we remove the PartialOrder$(\preceq)$
from $K(\varphi)$ and adjust DomainSubset$(u, w)$ and Persistent$(u, w)$, e.g. remove quantification over $w$ and the precondition $u\preceq w$ from $K(\varphi)$ and replace DomainSubset$(u, w)$ with
$$
	\bigwedge\{\forall u\forall x\forall\vec y(E(x, u)\to E(x, f_\psi(\vec y, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q\}
$$
and Persistent$(u, w)$ with
$$
	\bigwedge\{\forall\vec x\forall\vec y\forall u(A(\vec x, u)\to A(\vec x, f_\psi(\vec y, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q, \text{$A$ a predicate in $\varphi$}\}.
$$Putting all of this together yields the following translation:
\begin{theorem}\label{thm:fo-translation}
	Let $\mathcal R,\mathcal X,\mathcal Q$ be as above. For every $n$-ary predicate $A$ define a new $n+1$-ary predicate of the same name. For every $\psi\in\mathcal \mathcal X\cup\mathcal Q$ define a new function symbol $f_\psi$. Consider a new binary predicate symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$. Obtain $\mathcal R^\#$ by including
	$$
		\forall u\forall \vec x\left(\vec E(\vec x, u)\to\bigwedge_i A_i(\vec a_i, u)\to\bigvee_j B_j(\vec b_j, u)\right)
	$$
 	for each flat clause
 	$
 		\forall \vec x\left(\bigwedge_i A_i(\vec a_i)\to\bigvee_j B_j(\vec b_j)\right)\in\mathcal R,
 	$
 	$\mathcal X^\#$ by including
 	$$
 		\forall u\forall \vec x\left(A(\vec a, f_\psi(\vec x, u))\to B(\vec b, f_\psi(\vec x, u))\to C(\vec c, u)\right)
 	$$
 	for each implication clause
	$
		\psi = \forall \vec x\left(\left(A(\vec a)\to B(\vec b)\right)\to C(\vec c)\right)\in\mathcal X,
	$
	and $\mathcal Q^\#$ by including
	$$
		\forall u\forall \vec x\left(\forall y\left(A(\vec a, f_\psi(\vec x, u))\right)\to B(\vec b, u)\right)
	$$
	for each implication clause
	$
		\psi = \forall \vec x\left(\forall y\left(A(\vec a)\right)\to B(\vec b)\right)\in\mathcal X.
	$
	Then $\varphi$ is intuitionistically valid if and only if
	$$
		\varphi^\# := \exists xE(x, s)\to K(\varphi)\to \left(\bigwedge\mathcal R^\#\wedge\bigwedge\mathcal X^\#\wedge\bigwedge\mathcal Q^\#\right)\to P^\#(s)
	$$
	is classically valid, where $s$ is a new constant symbol and
	\begin{align*}
		K(\varphi) := &\:\bigwedge\left\{\forall\vec z\forall x\forall u((E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q\right\}\wedge\\
		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ a function in $\varphi$}\}\right)\wedge\\
		&\resizebox{.9\hsize}{!}{$\:\bigwedge\left\{\forall\vec z_1\forall \vec z_2\forall u(A(\vec z_1, u)\to A(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal X\cup\mathcal Q$, $A$ a predicate in $\varphi$}\right\}$.}
	\end{align*}
	The size of the obtained formula is linear in the input. However $|\mathcal X| + |\mathcal Q| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary predicate has been extended to a $n+1$-ary predicate.
\end{theorem}

It is not possible to completely eliminate worlds from the domain by enumerating them since there are formulas with no finite counter-examples. An analogous procedure to the propositional case fails, however we can still retrieve additional axioms that restrict the interpretation of functions. That is, we may assume that $f_\psi(\vec t, w) = w$, if $\exists u(w = f_\psi(\vec t, u))$ for all $\psi\in\mathcal X\cup\mathcal Q$ (i.e. that $f_\psi$ parametrized by $\vec t$ is idempotent), and $f_\psi(\vec t, w) = w$, if $\exists u(w \geq f_\psi(\vec t, u))$ for all $\psi\in\mathcal X$.

@@

\begin{wrapfigure}[5]{r}{0.4\textwidth}
	\vspace*{-1.6cm}
	\begin{tabular}{l|c|c}
		Embedding&571&21.4\%\\
		Embedding (idem.)&557&20.9\%\\
		ileanCoP 1.2&875&32.8\%\\
		nanoCoP-i 2.0&858&32.1\%\\\hline
		Total&2669&100\%
	\end{tabular}
	\caption{Solved problems.}
	\label{fig:a}
\end{wrapfigure}

@@

\begin{wrapfigure}[14]{r}{0.4\textwidth}
	\vspace{6.4cm}
	\hspace*{1.5cm}
	\begin{tikzpicture}
		\begin{scope}[opacity=0.1, blend mode=multiply]
			\def\firstcircle{(90:.75) circle (1.5)}
			\def\secondcircle{(210:.75) circle (1.5)}
			\def\thirdcircle{(330:.75) circle (1.5)}
			\begin{tikzpicture}
				\begin{scope}[shift={(3cm,-5cm)}, fill opacity=0.5]
					\fill[red] \firstcircle;
					\fill[green] \secondcircle;
					\fill[blue] \thirdcircle;
					\draw node {$491$};
					\draw node at (30:1.2) {$30$};
					\draw node at (150:1.2) {$42$};
					\draw node at (270:1.2) {$22$};
					\draw node at (90:1.8) {$8$};
					\draw node at (210:1.8) {$2$};
					\draw node at (330:1.8) {$417$};
				\end{scope}
			\end{tikzpicture}
		\end{scope}
	\end{tikzpicture}
	\vspace*{-7.6cm}
	\caption{Instances proven by combinations of Embedding (red), Embedding (idem.) (green), ileanCoP or nanoCoP (blue)}
	\label{fig:c}
\end{wrapfigure}

\vspace*{-.5cm}
Two versions of the translation were tested, one with and one without the additional axioms encoding idempotence. The results are stated in Figure \ref{fig:a}. 
In general, our approach falls short of the state of the art. 
However when grouping the benchmarks by $|\mathcal Q|$ and $|\mathcal X|$, see Figure~\ref{fig:b},
we observe good performance of our approach for $|\mathcal Q| = 0$ and $|\mathcal X| \leq 2$ resp. $|\mathcal X| \geq 6$.
Figure~\ref{fig:c} shows that our approach solved 52 problems not solved by the state-of-the-art provers. Notably, while there is big overlap in solved instances, ileanCoP and nanoCoP also have a significant symmetric difference.
\begin{figure}
	\centering
	\begin{tabular}{c|c|lr|lr|lr|lr|c}
		$|\mathcal Q|$&$|\mathcal X|$&\multicolumn{2}{c}{Embedding}&\multicolumn{2}{c}{Embedding (idem.)}&\multicolumn{2}{c}{ileanCoP}&\multicolumn{2}{c}{nanoCoP-i 2.0}&Total\\\hline
		0&0		&157&88.2\%	&157&88.2\%		&128&71.9\%		&61&38.1\%	&178\\
		0&1-2	&119&29.8\%&130&32.5\%		&108&27.0\%		&96&24.0\%	&400\\
		0&3+	&33&18.9\%	&38&21.7\%		&70&40.0\%		&51&29.1\%	&175\\
		1-2&0-2	&84&29.3\%	&81&28.2\%		&227&79.1\%		&221&77.0\%	&287\\
		1-2&3+	&102&24.7\%	&98&27.3\%		&180&43.6\%		&157&39.1\%	&413\\
		3+&0+	&66&5.4\%	&63&5.2\%		&262&20.7\%		&270&22.2\%	&1216\\
	\end{tabular}
	\caption{Solved problems grouped by $|\mathcal Q|$ and $|\mathcal X|$.}
	\label{fig:b}
\end{figure}

We conclude that our approach already has merits when $|\mathcal Q| = 0$ and $|\mathcal X|$ is small. 
We see that encoding idempotence does help to solve some additional benchmarks, but in general does not improve performance.
However, idempotence is a quite powerful concept from a model-theoretic perspective. We suspect that a more clever encoding of this property could improve results. There are other restrictions on models we could consider, based on structural properties of the formula, e.g. the relation between occurrences of different atoms.
